Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
revised abstract

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{This study analyzed highly correlated, feature-rich datasets from hyperspectral remote sensing data using multiple statistical and machine-learning methods.
The effect of filter-based feature selection methods on predictive performance was compared.
Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
Defoliation of trees (\%), derived from in-situ measurements from fall 2016, was modeled as a function of reflectance.
Variable importance was assessed using permutation-based feature importance.
Overall, the support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
For some combinations of learners and feature sets, filter methods achieved better predictive performances than using no feature selection.
Ensemble filters did not have a substantial impact on performance.
Permutation-based feature importance estimated features around the red edge to be most important.
However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the overall best performances.
More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.}

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{This study analyzed highly correlated, feature-rich datasets from hyperspectral remote sensing data using multiple statistical and machine-learning methods.
The effect of filter-based feature selection methods on predictive performance was compared.
Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
Defoliation of trees (\%), derived from in-situ measurements from fall 2016, was modeled as a function of reflectance.
Variable importance was assessed using permutation-based feature importance.
Overall, the support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), and lasso (L1) and ridge (L2) regressions by at least three percentage points.
The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
For some combinations of learners and feature sets, filter methods achieved better predictive performances than using no feature selection.
Ensemble filters did not have a substantial impact on performance.
The most important features were located around the red edge.
Additional features in the near-infrared region (800 nm - 1000 nm) were also essential to achieve the overall best performances.
Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.
Nevertheless, more training data and replication in similar benchmarking studies are needed to be able to generalize the results.}
